---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
##Problem 1
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#Using the head function to visually inspect the data 
head(NYC_Transit_Subway_Entrance_And_Exit_Data)
#From the head function output we can see that this data set contains 6 rows and 32 columns
names(NYC_Transit_Subway_Entrance_And_Exit_Data)
#cleaning the data to include only retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance

nyc_transit_cleaned <- NYC_Transit_Subway_Entrance_And_Exit_Data %>%
  select('Line', 'Station Name', 'Station Latitude', 'Station Longitude', 'Entry', 'Route1', 'Route2', 'Route3', 'Route4', 'Route5', 'Route6', 'Route7', 'Route8', 'Route9', 'Route10', 'Vending', 'Entrance Type', 'ADA' )

names(NYC_Transit_Subway_Entrance_And_Exit_Data)

#Convert the entry variable from character (YES vs NO) to a logical vector. 
nyc_transit_cleaned <- nyc_transit_cleaned %>%
mutate(Entry = ifelse(Entry == "YES", TRUE, FALSE))

#number of distinct stations


```
* The original data set had 6 rows and 32 columns and contained a large number of variables: "Division", "Line" ,"Station Name" ,"Station Latitude", "Station Longitude", "Route1"          "Route2", "Route3", "Route4", "Route5", "Route6", "Route7", "Route8", "Route9", "Route10",         "Route11", "Entrance Type", "Entry", "Exit Only", "Vending", "Staffing", "Staff Hours", "ADA"      "ADA Notes", "Free Crossover", "North South Street", "East West Street", "Corner",           
 "Entrance Latitude",  "Entrance Longitude", "Station Location",   "Entrance Location". The first step in our data cleaning process was to omit a number of variables that we are not interested in including in our analysis. After this cleaning step, the data set had 1868 observations over 18 variables.Next, I changed the Entry variable from its original YES/NO character state into a logical vector. A dataset is tidy if each column corresponds to a variable and each row corresponds to an observation. This data is tidy because each column represents a separate variable and each row is a unique observation. For example, here we can say that this cleaned data set has 1868 observations across 18 variables. 
 
``` {r setup, include=FALSE}
##Problem #2
library(tidyr)
library(readxl)
##omiting missing data
trash_data_cleaned <- X202409_Trash_Wheel_Collection_Data %>%
  filter(!is.na(Dumpster))

##rounding and then converting to an interger value. 
trash_data_cleaned <- trash_data_cleaned %>%
  rename(sports_balls = `Sports Balls`) %>%
  mutate(sports_balls = round(sports_balls),  
         sports_balls = as.integer(sports_balls)) 


## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:
 

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
